---
title: "Chapter 5 - Foundations for Inference"
author: ""
output:
    pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Heights of adults.** (7.7, p. 260) Researchers studying anthropometry collected body girth measurements and skeletal diameter measurements, as well as age, weight, height and gender, for 507 physically active individuals. The histogram below shows the sample distribution of heights in centimeters.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=3}
library(openintro)
data(bdims)
par(mar=c(3.7,2.5,0.5,0.5), las=1, mgp=c(2.5,0.7,0), cex.lab = 1.5)
histPlot(bdims$hgt, col = COL[1], xlab = "Height", ylab = "")
```

(a) What is the point estimate for the average height of active individuals? What about the median?

> The mean is 171.1 and the median is 170.3

```{r message=FALSE}
library(dplyr)
bdims %>%
  select(hgt) %>%
  summarise(avg = mean(hgt), median = median(hgt))
```


(b) What is the point estimate for the standard deviation of the heights of active individuals? What about the IQR?

> The standard deviation is 9.4 and the IQR is 14

```{r}
bdims %>%
  select(hgt) %>%
  summarise(sd = sd(hgt), IQR = IQR(hgt))
```

(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? And is a person who is 1m 55cm (155cm) considered unusually short? Explain your reasoning.

> For the person that is 180 cm it is not considered unusually tall because they are less than 1 sd away from the mean. On the otherhand the 155 cm person is considered unusually short because they are 1.71 sd away from the mean

```{r 180cm}
(180 - mean(bdims$hgt)) / sd(bdims$hgt)
```

```{r 155cm}
(155 - mean(bdims$hgt)) / sd(bdims$hgt)
```

(d) The researchers take another random sample of physically active individuals. Would you expect the mean and the standard deviation of this new sample to be the ones given above? Explain your reasoning.

> If the researcher takes another random sample the mean and the standard deviation is not going to be the same as there can be people that are very short or very tall whichi will change the mean and the standard deviation. What we hope is that the new sample is going to be close to the other sample and that the new mean and standard deviation will be close to each other

(e) The sample means obtained are point estimates for the mean height of all active individuals, if the sample of individuals is equivalent to a simple random sample. What measure do we use to quantify the variability of such an estimate (Hint: recall that $SD_x = \frac{\sigma}{\sqrt{n}}$)? Compute this quantity using the data from the original sample under the condition that the data are a simple random sample.

> The SE = 0.41

```{r}
sd(bdims$hgt)/sqrt(nrow(bdims))
```




--------------------------------------------------------------------------------

\clearpage

**Thanksgiving spending, Part I.** The 2009 holiday retail season, which kicked off on November 27, 2009 (the day after Thanksgiving), had been marked by somewhat lower self-reported consumer spending than was seen during the comparable period in 2008. To get an estimate of consumer spending, 436 randomly sampled American adults were surveyed. Daily consumer spending for the six-day period after Thanksgiving, spanning the Black Friday weekend and Cyber Monday, averaged $84.71. A 95% confidence interval based on this sample is ($80.31, $89.11). Determine whether the following statements are true or false, and explain your reasoning.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=3}
library(openintro)
data(data(thanksgiving_spend))
par(mar=c(3.7,2.2,0.5,0.5), las=1, mgp=c(2.5,0.7,0), cex.lab = 1.5)
histPlot(thanksgiving_spend$spending, col = COL[1], xlab = "Spending", ylab = "")
```

(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and $89.11.

> FALSE - The confidence interval can only be applied to population and not samples

(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.

> FALSE - the confidence interval can still be used even if it is skewed as long as the sample size is large enough

(c) 95% of random samples have a sample mean between $80.31 and $89.11.

> FALSE - if we ran the sample again we may get a different confidence interval and may not fall in the confidence interval

(d) We are 95% confident that the average spending of all American adults is between $80.31 and $89.11.

> TRUE - based on the sample we are able to state that we are confident that 95% of American adults will spend inside the confidence interval

(e) A 90% confidence interval would be narrower than the 95% confidence interval since we don't need to be as sure about our estimate.

> TRUE - When we decrease the confidence it becomes narrower but we are also less confident

(f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is now, we would need to use a sample 3 times larger.

> FALSE - We will need to increase the sample size by 9 times

\begin{align}
3 * SE = \frac{sigma}{\sqrt{n}} \\
SE = \frac{sigma}{3 * \sqrt{n}} \\
SE = \frac{sigma}{\sqrt{3^2 * n}} \\
SE = \frac{sigma}{\sqrt{9 * n}} \\
\end{align}

(g) The margin of error is 4.4.

> TRUE

```{r}
(89.11 - 80.31) / 2
```





--------------------------------------------------------------------------------

\clearpage

**Gifted children, Part I.** Researchers investigating characteristics of gifted children col- lected data from schools in a large city on a random sample of thirty-six children who were identified as gifted children soon after they reached the age of four. The following histogram shows the dis- tribution of the ages (in months) at which these children first counted to 10 successfully. Also provided are some sample statistics.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=3}
library(openintro)
data(gifted)
par(mar=c(3.7,2.2,0.5,0.5), las=1, mgp=c(2.5,0.7,0), cex.lab = 1.5)
histPlot(gifted$count, col = COL[1], 
         xlab = "Age child first counted to 10 (in months)", ylab = "", 
         axes = FALSE)
axis(1)
axis(2, at = c(0,3,6))
```

\begin{tabular}{r | l}
n   & 36 \\
min & 21 \\
mean    & 30.69 \\
sd  & 4.31 \\
max & 39 
\end{tabular}

(a) Are conditions for inference satisfied?

> Yes the children are randomly selected and the random sample is larger than 30 samples

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children fist count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10.

> Null Hypothesis = 32 months | Alternative hypothesis < 32 months | Significance = 0.10

(c) Interpret the p-value in context of the hypothesis test and the data.

```{r}
1 - pnorm(32, mean=30.69, sd = 4.31/sqrt(36))
```
> Because the p value 0.03 is less than 0.10 we will be rejecting the null hypothesis and instead will be accepting the alternative hypothesis

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.

```{r}
z = 1.645
mean = 30.69
sd = 4.31
n = 36
SE = sd / sqrt(n)

high = mean + z * SE
low = mean - z * SE

high
low
```

> The 90% confidence intercal is (29.5, 31.9)

(e) Do your results from the hypothesis test and the confidence interval agree? Explain.

> Yes because the 90% confidence interval is (29.5, 31.9) and we accepted the alternative hypothesis



--------------------------------------------------------------------------------

\clearpage

**Gifted children, Part II.** Exercise above describes a study on gifted children. In this study, along with variables on the children, the researchers also collected data on the mother's and father's IQ of the 36 randomly sampled gifted children. The histogram below shows the distribution of mother's IQ. Also provided are some sample statistics.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=3}
library(openintro)
data(gifted)
par(mar=c(3.7,2.2,0.5,0.5), las=1, mgp=c(2.5,0.7,0), cex.lab = 1.5)
histPlot(gifted$motheriq, col = COL[1], 
         xlab = "Mother's IQ", ylab = "", axes = FALSE)
axis(1)
axis(2, at = c(0,4,8,12))
```

\begin{tabular}{r | l}
n   & 36 \\
min & 101 \\
mean    & 118.2 \\
sd  & 6.5 \\
max & 131 
\end{tabular}

(a) Perform a hypothesis test to evaluate if these data provide convincing evidence that the average IQ of mothers of gifted children is different than the average IQ for the population at large, which is 100. Use a significance level of 0.10.

> Null Hypothesis = 100 IQ | Alternative Hypothesis > 100 IQ | Significance = 0.10

```{r}
mean = 118.2
sd = 6.5
n = 36
z = (mean - 100) / sd
1 - pnorm(z, mean = 0, sd = 1)
```

> We can see that 0.002 < 0.10 therefore we reject the null hypthesis and accept the alternative hypothesis

(b) Calculate a 90% confidence interval for the average IQ of mothers of gifted children.

```{R}
mean = 118.2
sd = 6.5
n = 36
z = 1.65

SE = sd / sqrt(n)

high = mean + z * SE
low = mean - z * SE

high
low
```

> The confidence interval is (116, 120)

(c) Do your results from the hypothesis test and the confidence interval agree? Explain.

> Yes because we rejected the null hypothesis which stated that the average IQ is 100 but is not in the confidence interval. 





--------------------------------------------------------------------------------

\clearpage

**CLT.** Define the term "sampling distribution" of the mean, and describe how the shape, center, and spread of the sampling distribution of the mean change as sample size increases.

> when the sample size increases
> - Becomes closer to normal distribution
> - The center of the normal distribution becomes taller
> - The normal distribution spread becomes smaller



--------------------------------------------------------------------------------

\clearpage

**CFLBs.** A manufacturer of compact fluorescent light bulbs advertises that the distribution of the lifespans of these light bulbs is nearly normal with a mean of 9,000 hours and a standard deviation of 1,000 hours.

(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours?

> The probablilty of the light bulb lasting more than 10,500 hours is 6.6%

```{r}
mean = 9000
sd = 1000
1 - pnorm(10500, mean = mean, sd = sd)
```

(b) Describe the distribution of the mean lifespan of 15 light bulbs.

> The distribution of the 15 light bulbs will follow a normal distribution

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?

> The probablity of mean of 15 light bulbs lasting more than 10,500 is 0%

```{r}
mean = 9000
sd = 1000
n = 15

1 - pnorm(10500, mean = 9000, sd = 1000/sqrt(10))
```


(d) Sketch the two distributions (population and sampling) on the same scale.

```{r}
set.seed(1)
library(ggplot2)

sample = rnorm(15, mean=9000, sd = 1000)
population = rnorm(1500, mean = 9000, sd = 1000)

ggplot() + aes(sample) + geom_histogram()
ggplot() + aes(population) + geom_histogram()
```

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution?

> No as we assumed that the distribution was normal and the pnorm function requires a normal distribution





--------------------------------------------------------------------------------

\clearpage

**Same observation, different sample size.** Suppose you conduct a hypothesis test based on a sample where the sample size is n = 50, and arrive at a p-value of 0.08. You then refer back to your notes and discover that you made a careless mistake, the sample size should have been n = 500. Will your p-value increase, decrease, or stay the same? Explain.

> In the calcuation for standard error the sample size is in the denominator. This means that as the sample size increases the SE will decrease. This means that the P value will also decrease





